Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-01-07 manuscript/manuscript.Rmd
updated with supp mat

@@

Figures
-------

```{r figure_1, dependson="plot-options", fig.cap="The proportion of interannual variability in  vital rates explained by the climate covariates. The contribution for growth is defined as: (Climate model - Constant Model)/(Full model - Constant model). The contribution for survival and colonization, where we could not estimate a full model with year random effects at the quadrat level, is defined as: (Constant Model - Climate Model)/Constant Model.", fig.width=7, fig.height=3, cache=FALSE, results='hide'}
deviance <- data.frame(VitalRate = c("Growth", "Survival", "Colonization"),
                       ConstantModel = rep(NA,3),
                       ClimateModel = rep(NA,3),
                       FullModel = rep(NA,3),
                       ClimateContribution = rep(NA,3))
deviance$FullModel[1] <- round(read.csv("../montana/quadBM/vitalRateRegressions/growth/growthDeviance.csv")[,2],2)
deviance$ClimateModel[1] <- round(read.csv("../montana/quadBM/vitalRateRegressions/growth/growthDevianceCLIMATE.csv")[,2],2)
deviance$ConstantModel[1] <- round(read.csv("../montana/quadBM/vitalRateRegressions/growth/growthDevianceCONSTANT.csv")[,2],2)
deviance$ClimateContribution[1] <- with(deviance[1,], (ClimateModel-ConstantModel)/(FullModel-ConstantModel))
deviance$ClimateModel[2] <- round(read.csv("../montana/quadBM/vitalRateRegressions/survival/survivalDeviance.csv")[,2],2)
deviance$ConstantModel[2] <- round(read.csv("../montana/quadBM/vitalRateRegressions/survival/survivalDevianceCONSTANT.csv")[,2],2)
deviance$ClimateModel[3] <- round(read.csv("../montana/quadBM/vitalRateRegressions/colonization/colonizationDeviance.csv")[,2],2)
deviance$ConstantModel[3] <- round(read.csv("../montana/quadBM/vitalRateRegressions/colonization/colonizationDevianceCONSTANT.csv")[,2],2)
deviance$ClimateContribution[2:3] <- with(deviance[2:3,], (ConstantModel-ClimateModel)/(ConstantModel))

colnames(deviance) <- c("Vital Rate", "Constant model", "Climate model", "Full model", "Contribution of climate covariates")
library(ggplot2)
devGQ <- data.frame(VitalRate = c("Growth", "Survival", "Colonization"),
                       ConstantModel = rep(NA,3),
                       ClimateModel = rep(NA,3),
                       FullModel = rep(NA,3),
                       ClimateContribution = rep(NA,3))
devGQ$FullModel[1] <- round(read.csv("../montana/ipm/vitalRateRegs/growth/growthDeviance.csv")[,2],2)
devGQ$ClimateModel[1] <- round(read.csv("../montana/ipm/vitalRateRegs/growth/growthDevianceCLIMATE.csv")[,2],2)
devGQ$ConstantModel[1] <- round(read.csv("../montana/ipm/vitalRateRegs/growth/growthDevianceCONSTANT.csv")[,2],2)
devGQ$FullModel[2] <- round(read.csv("../montana/ipm/vitalRateRegs/survival/survivalDeviance.csv")[,2],2)
devGQ$ClimateModel[2] <- round(read.csv("../montana/ipm/vitalRateRegs/survival/survivalDevianceCLIMATE.csv")[,2],2)
devGQ$ConstantModel[2] <- round(read.csv("../montana/ipm/vitalRateRegs/survival/survivalDevianceCONSTANT.csv")[,2],2)
devGQ$FullModel[3] <- round(read.csv("../montana/ipm/vitalRateRegs/recruitment/recruitmentDeviance.csv")[,2],2)
devGQ$ClimateModel[3] <- round(read.csv("../montana/ipm/vitalRateRegs/recruitment/recruitmentDevianceCLIMATE.csv")[,2],2)
devGQ$ConstantModel[3] <- round(read.csv("../montana/ipm/vitalRateRegs/recruitment/recruitmentDevianceCONSTANT.csv")[,2],2)
devGQ$ClimateContribution[1] <- with(devGQ[1,], (ClimateModel-ConstantModel)/(FullModel-ConstantModel))
devGQ$ClimateContribution[2] <- with(devGQ[2,], (ConstantModel-ClimateModel)/(ConstantModel))
devGQ$ClimateContribution[3] <- with(devGQ[3,], (ConstantModel-ClimateModel)/(ConstantModel))

devM <- devGQ[,c(1,5)]
devM$Level <- rep("Individual", 3)
tmpD <- deviance[,c(1,5)]
tmpD$Level <- rep("Quadrat", 3)
colnames(tmpD)[2] <- "value"
colnames(tmpD)[1] <- "VitalRate"
colnames(devM)[2] <- "value"
plotD <- rbind(tmpD, devM)
gClimCont1 <- ggplot(data=subset(plotD, Level=="Individual"), aes(x=VitalRate, y=value))+
  geom_bar(stat="identity", position="dodge", fill="grey30")+
  ylab("Contribution of Climate Effects")+
  xlab("Vital Rate")+
  ggtitle("Individual level")+
  scale_y_continuous(limits=c(0, 0.5))+
  scale_x_discrete(labels=c("Recruitment", "Growth", "Survival"))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 25, hjust = 1))
gClimCont2 <- ggplot(data=subset(plotD, Level=="Quadrat"), aes(x=VitalRate, y=value))+
  geom_bar(stat="identity", position="dodge", fill="grey30")+
  ylab("Contribution of Climate Effects")+
  xlab("Vital Rate")+
  ggtitle("Quadrat level")+
  scale_y_continuous(limits=c(0, 0.5))+
  scale_x_discrete(labels=c("Colonization", "Cover change", "Extirpation"))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 25, hjust = 1))
library(gridExtra)
gClim <- grid.arrange(gClimCont1, gClimCont2, ncol=2, nrow=1)
gClim

```{r figure_2, dependson="plot-options", fig.cap="Posterior means (vertical ticks), 75% credible intervals (heavy lines), and 95% credible intervals (light lines) of climate effects on growth at both levels of inferences. The dashed vertical line is at 0, indicating no effect. Horizontal line at 0 indicates perfect agreement between mean observed cover in that year and the model predictions.", fig.width=3, fig.height=7, cache=FALSE, results='hide'}
gQC <- read.csv("../montana/quadBM/vitalRateRegressions/growth/growthQuants.csv")
gQC <- gQC[129:144,]
coefNames <- c("ppt1", "ppt2", "temp1", "temp2")
gQC$X <- rep(coefNames, each=4)
gQC$rate <- "growth"
sQC <- read.csv("../montana/quadBM/vitalRateRegressions/survival/survivalQuants.csv")
sQC <- sQC[33:48,]
coefNames <- c("ppt1", "ppt2", "temp1", "temp2")
sQC$X <- rep(coefNames, each=4)
sQC$rate <- "survival"
rQC <- read.csv("../montana/quadBM/vitalRateRegressions/colonization/colonizationQuants.csv")
rQC <- rQC[29:44,]
coefNames <- c("ppt1", "ppt2", "temp1", "temp2")
rQC$X <- rep(coefNames, each=4)
rQC$rate <- "colonization"
plotQC <- rbind(gQC,sQC,rQC)
gQ <- ggplot(plotQC, aes(x=species, y=X50., color=rate))+
  geom_hline(aes(yintercept=0), linetype="dashed")+
  geom_errorbar(aes(ymin=X2.5., ymax=X97.5.), width=0, position=position_dodge(width=0.5))+
  geom_errorbar(aes(ymin=X25., ymax=X75.), width=0, size=1.25, position=position_dodge(width=0.5))+
  geom_point(size=4, position=position_dodge(width=0.5), shape="|")+
  ylab("Standardized Coefficient Value")+
  xlab("Species")+
  facet_grid(X~.)+
  coord_flip()+
  theme_bw()+
  scale_color_manual(values=c("grey25", "steelblue", "slateblue"))+
  guides(color=guide_legend(title=""))+
  theme(legend.position="top")+
  ggtitle("Quadrat level")

gIC <- read.csv("../montana/ipm/vitalRateRegs/growth/growthQuants.csv")
gIC <- gIC[133:148,]
coefNames <- c("ppt1", "ppt2", "temp1", "temp2")
gIC$X <- rep(coefNames, each=4)
gIC$rate <- "growth"
sIC <- read.csv("../montana/ipm/vitalRateRegs/survival/survivalQuants.csv")
sIC <- sIC[133:148,]
coefNames <- c("ppt1", "ppt2", "temp1", "temp2")
sIC$X <- rep(coefNames, each=4)
sIC$rate <- "survival"
rIC <- read.csv("../montana/ipm/vitalRateRegs/recruitment/recruitmentQuants.csv")
rIC <- rIC[81:96,]
coefNames <- c("ppt1", "ppt2", "temp1", "temp2")
rIC$X <- rep(coefNames, each=4)
rIC$rate <- "colonization"
plotIC <- rbind(gIC,sIC,rIC)
gI <- ggplot(plotIC, aes(x=species, y=X50., color=rate))+
  geom_hline(aes(yintercept=0), linetype="dashed")+
  geom_errorbar(aes(ymin=X2.5., ymax=X97.5.), width=0, position=position_dodge(width=0.5))+
  geom_errorbar(aes(ymin=X25., ymax=X75.), width=0, size=1.25, position=position_dodge(width=0.5))+
  geom_point(size=4, position=position_dodge(width=0.5), shape="|")+
  ylab("Standardized Coefficient Value")+
  xlab("Species")+
  facet_wrap(X~.)+
  coord_flip()+
  theme_bw()+
  scale_color_manual(values=c("grey25", "steelblue", "slateblue"),
                     labels=c("recruitment", "growth", "survival"))+
#   guides(color=FALSE)+
  guides(color=guide_legend(title=""))+
  theme(legend.position="top")+
  ggtitle("Individual level")

####
#### Plot just for growth
####
growthIC <- subset(plotIC, rate=="growth")
growthQC <- subset(plotQC, rate=="growth")
growthIC$Level <- "Individual"
growthQC$Level <- "Quadrat"
growthAll <- rbind(growthIC, growthQC)

gGrowth <- ggplot(growthAll, aes(x=species, y=X50., color=Level))+
  geom_hline(aes(yintercept=0), linetype="dashed")+
  geom_errorbar(aes(ymin=X2.5., ymax=X97.5.), width=0, position=position_dodge(width=0.5))+
  geom_errorbar(aes(ymin=X25., ymax=X75.), width=0, size=1.25, position=position_dodge(width=0.5))+
  geom_point(size=4, position=position_dodge(width=0.5), shape="|")+
  ylab("Standardized Coefficient Value")+
  xlab("Species")+
  facet_grid(X~.)+
  coord_flip()+
  theme_bw()+
  scale_color_manual(values=c("grey25", "slateblue"))+
#   guides(color=FALSE)+
  guides(color=guide_legend(title=""))+
  theme(legend.position="top")
print(gGrowth)
```

```{r figure_3, dependson="plot-options", fig.cap="Boxplots of model residuals for one-step-ahead forecasts at each observation year. Each one-step forecast was simulated `r nSims` times. Note that the y-axes vary across panels. The light blue line shows the difference between the observed-year percent cover and the average cover observed across all years. The models tend to underpredict and perform poorly when observed cover in a given year is a large deviant from the mean.", fig.width=5, fig.height=6, cache=FALSE, results='hide'}
library(plyr)
library(reshape2)
resD <- readRDS("../montana/quadBM/simulations/quadBM_oneStep_Residuals.rds")
nSims <- length(unique(resD$sim))
diffD <- ddply(resD, c("year", "species"), summarise,
               avgDiff = mean(yearDiff))

ggplot(resD)+
  geom_hline(aes(yintercept=0))+
#   geom_line(data=diffD, aes(x=year, y=avgDiff, group=species), color="purple", alpha=0.5)+
  stat_smooth(data=diffD, aes(x=year, y=avgDiff, group=species), color="lightblue", se = FALSE, size=1)+
  geom_boxplot(aes(x=year, y=covChangeResiduals), outlier.shape = 1, width=0.3, outlier.size = 0.5)+
  facet_grid(species~., scales = "free_y")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r figure_4, dependson="plot-options", fig.cap="Simulated changes species' median cover caused by a 1% increase in observed precipitation or temperature. We used the median of simulated results because the average was highly influence by some extreme outliers.", fig.width=4.5, fig.height=4, cache=FALSE, results='hide'}
climBase <- readRDS("../montana/quadBM/simulations/climateChangeBase.rds")
climPpt <- readRDS("../montana/quadBM/simulations/climateChangePrecip.rds")
climTemp <- readRDS("../montana/quadBM/simulations/climateChangeTemperature.rds")
climBoth <- readRDS("../montana/quadBM/simulations/climateChangePptTmp.rds")
climBase$sim <- rep("Observed", nrow(climBase)) 
climPpt$sim <- rep("Precipitation", nrow(climBase)) 
climTemp$sim <- rep("Temperature", nrow(climBase)) 
climBoth$sim <- rep("Both", nrow(climBase)) 
climSimsQBM <- rbind(climBase, climPpt, climTemp, climBoth)
# ggplot(climSimsQBM)+
#   geom_boxplot(aes(x=species, y=cover*100, fill=sim))+
#   scale_y_continuous(limits=c(0,10))
climSimsQBMavg <- ddply(climSimsQBM, .(sim, species), summarise,
                        avgCover = median(cover))
climSimQBMcast <- dcast(climSimsQBMavg, formula = species~sim)
climSimQBMcast$diffP <- with(climSimQBMcast, (Precipitation*100)-(Observed*100)) 
climSimQBMcast$diffT <- with(climSimQBMcast, (Temperature*100)-(Observed*100))
climSimQBMcast$diffB <- with(climSimQBMcast, (Both*100)-(Observed*100))
climSimQBMplot <- melt(climSimQBMcast)
ggplot(data=subset(climSimQBMplot, variable=="diffP"|variable=="diffT"|variable=="diffB"))+
  geom_bar(aes(x=species, y=value, fill=variable), stat="identity", position="dodge", color="white")+
  scale_fill_manual(name="", labels=c("1% Precipitation Increase", "1% Temperature Increase", "1% Temp. & Ppt. Increase"), values=c("grey25", "grey50", "grey75"))+
  theme_bw()+
  theme(legend.position=c(0.6, 0.8))+
  ylab("Cover Change (%)")+
  xlab("Species")
```

>>>>

```{r libraris, include=FALSE}
library(ggplot2)
library(plyr)
library(gridExtra)
library(reshape2)
```

@@

Figures
-------
```{r figure_1, dependson="plot-options", fig.cap="Boxplots of model residuals for one-step-ahead forecasts at each observation year. Each one-step forecast was simulated `r nSims` times. Note that the y-axes vary across panels. The light blue line shows the difference between the observed-year percent cover and the average cover observed across all years. The models tend to underpredict and perform poorly when observed cover in a given year is a large deviant from the mean.", fig.width=5, fig.height=6, cache=FALSE, results='hide'}
resD <- readRDS("../montana/quadBM/simulations/quadBM_oneStep_Residuals.rds")
nSims <- length(unique(resD$sim))
diffD <- ddply(resD, c("year", "species"), summarise,
               avgDiff = mean(yearDiff))

ggplot(resD)+
  geom_hline(aes(yintercept=0))+
#   geom_line(data=diffD, aes(x=year, y=avgDiff, group=species), color="purple", alpha=0.5)+
#   stat_smooth(data=diffD, aes(x=year, y=avgDiff, group=species), color="lightblue", se = FALSE, size=1)+
  geom_boxplot(aes(x=year, y=covChangeResiduals), outlier.shape = 1, width=0.3, outlier.size = 0.5)+
  facet_grid(species~., scales = "free_y")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r figure_2, dependson="plot-options", fig.cap="Simulated changes species' median cover caused by a 1% increase in observed precipitation or temperature. We used the median of simulated results because the average was highly influence by some extreme outliers.", fig.width=4.5, fig.height=4, cache=FALSE, results='hide'}
climBase <- readRDS("../montana/quadBM/simulations/climateChangeBase.rds")
climPpt <- readRDS("../montana/quadBM/simulations/climateChangePrecip.rds")
climTemp <- readRDS("../montana/quadBM/simulations/climateChangeTemperature.rds")
climBoth <- readRDS("../montana/quadBM/simulations/climateChangePptTmp.rds")
climBase$sim <- rep("Observed", nrow(climBase)) 
climPpt$sim <- rep("Precipitation", nrow(climBase)) 
climTemp$sim <- rep("Temperature", nrow(climBase)) 
climBoth$sim <- rep("Both", nrow(climBase)) 
climSimsQBM <- rbind(climBase, climPpt, climTemp, climBoth)
# ggplot(climSimsQBM)+
#   geom_boxplot(aes(x=species, y=cover*100, fill=sim))+
#   scale_y_continuous(limits=c(0,10))
climSimsQBMavg <- ddply(climSimsQBM, .(sim, species), summarise,
                        avgCover = median(cover))
climSimQBMcast <- dcast(climSimsQBMavg, formula = species~sim)
climSimQBMcast$diffP <- with(climSimQBMcast, (Precipitation*100)-(Observed*100)) 
climSimQBMcast$diffT <- with(climSimQBMcast, (Temperature*100)-(Observed*100))
climSimQBMcast$diffB <- with(climSimQBMcast, (Both*100)-(Observed*100))
climSimQBMplot <- melt(climSimQBMcast)
ggplot(data=subset(climSimQBMplot, variable=="diffP"|variable=="diffT"|variable=="diffB"))+
  geom_bar(aes(x=species, y=value, fill=variable), stat="identity", position="dodge", color="white")+
  scale_fill_manual(name="", labels=c("1% Precipitation Increase", "1% Temperature Increase", "1% Temp. & Ppt. Increase"), values=c("grey25", "grey50", "grey75"))+
  theme_bw()+
  theme(legend.position=c(0.6, 0.8))+
  ylab("Cover Change (%)")+
  xlab("Species")
```

### 2015-01-07 manuscript/manuscript.tex
updated bibliography with correct R citation
updated with supp mat

All of our analyses (model fitting and simulating) were conducted in R
(Team 2013). We used the MCMC sampler in JAGS (Plummer 2003) to estimate
the posterior distributions of model parameters and the package `r2jags'
(Su and Yajima 2012) to connect R to JAGS. We obtained posterior
distributions for all model parameters from three parallel MCMC chains
run for 50,000 iterations, after discarding an initial 50,000
iterations. We assessed convergence visually and using the Gelman and
Rubin (1992) diagnostic in the R package `coda' (Plummer et al. 2006).
Scale reduction factors for all parameters were less than 1.02,
indicating convergence. For the purposes of introducing stochasticity in
our population models, we saved the final 1,000 iterations from each
chain for all parameters to be used as randomly drawn values during
population simulation.

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_1.pdf}
\caption{The proportion of interannual variability in vital rates
explained by the climate covariates. The contribution for growth is
defined as: (Climate model - Constant Model)/(Full model - Constant
model). The contribution for survival and colonization, where we could
not estimate a full model with year random effects at the quadrat level,
is defined as: (Constant Model - Climate Model)/Constant Model.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_2.pdf}
\caption{Posterior means (vertical ticks), 75\% credible intervals
(heavy lines), and 95\% credible intervals (light lines) of climate
effects on growth at both levels of inferences. The dashed vertical line
is at 0, indicating no effect. Horizontal line at 0 indicates perfect
agreement between mean observed cover in that year and the model
predictions.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_3.pdf}
\caption{Boxplots of model residuals for one-step-ahead forecasts at
each observation year. Each one-step forecast was simulated
\texttt{r nSims} times. Note that the y-axes vary across panels. The
light blue line shows the difference between the observed-year percent
cover and the average cover observed across all years. The models tend
to underpredict and perform poorly when observed cover in a given year
is a large deviant from the mean.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_4.pdf}
\caption{Simulated changes species' median cover caused by a 1\%
increase in observed precipitation or temperature. We used the median of
simulated results because the average was highly influence by some
extreme outliers.}
\end{figure}

@@

@@

Team, R. 2013. R Development Core Team. R: A Language and Environment
for Statistical Computing.

>>>>

All of our analyses (model fitting and simulating) were conducted in R
(R Development Core Team 2013). We used the MCMC sampler in JAGS
(Plummer 2003) to estimate the posterior distributions of model
parameters and the package `r2jags' (Su and Yajima 2012) to connect R to
JAGS. We obtained posterior distributions for all model parameters from
three parallel MCMC chains run for 50,000 iterations, after discarding
an initial 50,000 iterations. We assessed convergence visually and using
the Gelman and Rubin (1992) diagnostic in the R package `coda' (Plummer
et al. 2006). Scale reduction factors for all parameters were less than
1.02, indicating convergence. For the purposes of introducing
stochasticity in our population models, we saved the final 1,000
iterations from each chain for all parameters to be used as randomly
drawn values during population simulation.

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_1.pdf}
\caption{Boxplots of model residuals for one-step-ahead forecasts at
each observation year. Each one-step forecast was simulated
\texttt{r nSims} times. Note that the y-axes vary across panels. The
light blue line shows the difference between the observed-year percent
cover and the average cover observed across all years. The models tend
to underpredict and perform poorly when observed cover in a given year
is a large deviant from the mean.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_2.pdf}
\caption{Simulated changes species' median cover caused by a 1\%
increase in observed precipitation or temperature. We used the median of
simulated results because the average was highly influence by some
extreme outliers.}
\end{figure}

@@

R Development Core Team. 2013. R Development Core Team. R: A Language
and Environment for Statistical Computing.

@@
